Gradual Soundness: Lessons from Static Python

### 2021-12-10 pj-7.1/paper.tex
pj: cleanup

%% Thesis:
%%  the addition of sound nominal types to an optionally-typed codebase
%%  can effectively improve peformance

@@

\begin{itemize}
  \item
    Evidence that gradual soundness delivers Python-level performance at a low migration cost
    and faster performance as programmers gradually incorporate sound types.
    \SP{} can be applied to a codebase with no refactoring and marginal performance overhead.
    After refactoring to sound types, \SP{} performs well on a microbenchmark
    suite and in production.
  \item
    A formal description of the \SP{} core language and an analysis of its soundness
    and optimizations.
    Soundness is gradual in the sense that some types are fully sound and others validate
    only top-level shapes.
  \item
    Discussions of the expressiveness limitations in \SP{} relative to
    idealized gradual type systems, and the programming discipline that Instagram
    follows in their gradually-sound codebase.
\end{itemize}

@@

Interesting, because StrongScript disallowed.
(But Nom allowed right?! Maybe allowed in a funny way, possibly to support their Dyn.)

@@

Surprisingly Unions are static only.
Except for unions to false because those are so common.

@@

%    No, there’s no overload resolution by arguments in Python. This difference
%    in behavior isn’t so much desired as just a consequence of making
%    `INVOKE_METHOD` optimizable. With a normal dynamic `CALL_FUNCTION`, first
%    the callable is placed on the stack, then the arguments, then there is a
%    `CALL_FUNCTION` (with number of args in oparg) to perform the call. This
%    means that first the callable is resolved, then the arguments. But with
%    `INVOKE_METHOD` we want to resolve the callable as part of the invoke
%    itself, since this gives us opportunity to inline-cache the target of the
%    call instead of always having to call something dynamic and unknown that’s
%    on the stack. So that necessarily implies that first the arguments are
%    resolved and placed on the stack, then the callable is resolved as part of
%    the `INVOKE_METHOD`. -Carl

%    2021-11-15: Dino says the JIT will check/upgrade INVOKE_METHOD to INVOKE_FUNCTION
%    2021-07-23: final classes get INVOKE_FUNCTION, not INVOKE_METHOD

@@

\begin{center}
  \shorturl{https://}{github.com/facebookincubator/cinder/issues/created\_by/LuKC1024}
\end{center}

\section{Under the Hood: Static Python Specifics}
\label{s:impl}

@@

%First-class classes are untyped
%(good discussion here:
%- mixin based tends to structrural types
%- nominality overly restrictive, get forced into <: hierarchy
%- but don't have NO IDEA for wat to do
%  + classes and mixins paper shows one idea,
%    ... need complicated mixin study??? (many sub objects)
%    built new type system to be mixin-aware, layered atop java
%    unsure if such extension works for SP
%    furthermor, class-and-mixins depends on Java single inheritance
%  + bracha cook mixin pattern
%  + tate muehlboeck structural object + interfaces

@@

CheckedDict enables strong type checks and optimizations, but is incredibly painful to use.
It is painful even though ChkDict supports the full dict API
(any context that uses a Python dict can use a ChkDict without code changes).

>>>>

%% Thesis:
%%  1. Static Python is sound, fast, and practical
%%  2. SP has unique features and restrictions for speed
%%  3. Gradual soundness is a feature: start with weak types and eventually migrate to checked types

@@

\begin{itemize}
  \item
    Evidence that gradual soundness delivers Python-level performance at a low migration cost
    and faster performance as programmers gradually incorporate sound types.
    \SP{} can be applied to a codebase with no refactoring and marginal performance overhead.
    After refactoring to sound types, \SP{} performs well on a microbenchmark
    suite and in production.
  \item
    A formal description of the \SP{} core language and an analysis of its soundness
    and optimizations.
    Soundness is gradual in the sense that some types are fully sound and others validate
    only top-level shapes.
  \item
    Discussions of the expressiveness limitations in \SP{} relative to
    idealized gradual type systems, and of the programming discipline that Instagram
    follows in their gradually-sound codebase.
\end{itemize}

@@

Interesting because StrongScript disallowed.
(But Nom allowed right?! Maybe allowed in a funny way, possibly to support their Dyn.)

@@

Unions are static only,
except for unions to false because those are common and can be implemented efficiently.

@@

%    No, there’s no overload resolution by arguments in Python. This difference
%    in behavior isn’t so much desired as just a consequence of making
%    `INVOKE_METHOD` optimizable. With a normal dynamic `CALL_FUNCTION`, first
%    the callable is placed on the stack, then the arguments, then there is a
%    `CALL_FUNCTION` (with number of args in oparg) to perform the call. This
%    means that first the callable is resolved, then the arguments. But with
%    `INVOKE_METHOD` we want to resolve the callable as part of the invoke
%    itself, since this gives us opportunity to inline-cache the target of the
%    call instead of always having to call something dynamic and unknown that’s
%    on the stack. So that necessarily implies that first the arguments are
%    resolved and placed on the stack, then the callable is resolved as part of
%    the `INVOKE_METHOD`. -Carl
%
%    2021-11-15: Dino says the JIT will check/upgrade INVOKE_METHOD to INVOKE_FUNCTION
%    2021-07-23: final classes get INVOKE_FUNCTION, not INVOKE_METHOD

@@

\begin{center}
  %% most filed by KC, a few by Ben
  \shorturl{https://}{github.com/facebookincubator/cinder/issues/created\_by/LuKC1024}
\end{center}

\section{Scaling to Python}
\label{s:impl}

@@

%Notes on first-class classes being untyped:
%- mixin-based code tends to structrural types
%- nominality overly restrictive, get forced into <: hierarchy
%- but don't have NO IDEA for wat to do
%  + classes and mixins paper shows one idea,
%    ... need complicated mixin study??? (many sub objects)
%    built new type system to be mixin-aware, layered atop java
%    unsure if such extension works for SP
%    furthermore, class-and-mixins depends on Java single inheritance
%  + bracha cook mixin pattern
%  + tate muehlboeck structural object + interfaces

@@

CheckedDict enables strong type checks and optimizations, but is painful to use.
It is painful even though ChkDict supports the full dict API
(any context that uses a Python dict can use a ChkDict without code changes).
